Stat 400 Presentation

### 2021-12-08 Stat 400 presentation.Rmd
Added in the plot, this code can probably be reduced

>>>>

## Plot of infected people

\tiny
```{r, echo = FALSE, output = TRUE}
library(ggplot2)
starting_vals <- function(num_subjects, I_0, x_dim, y_dim){

  return(df)
}

#Random step length function
get_step_length <- function(df, x_dim, y_dim){
  step_lengths <- rnorm(n=nrow(df), mean=min(x_dim, y_dim)/4, sd=min(x_dim, y_dim)/12)
  return(step_lengths)
}

#Random step direction function
get_step_direction <- function(length_vec){
  direction <- runif(n=length(length_vec), min=0, max=2*pi)
}

#Get delta_x, delta_y
get_delts <- function(length_vec, direc_vec){
  delta_x <- length_vec * cos(direc_vec)
  delta_y <- length_vec * sin(direc_vec)
  return(cbind(delta_x, delta_y))
}

#Function for getting updated coordinates
update_coords <- function(og_df, delta_df){
  new_mx <- matrix(data=NA, nrow=nrow(og_df), ncol=5)
  new_df <- data.frame(new_mx)
  new_df[,1] <- as.numeric(og_df[,'x']) + delta_df[,'delta_x']
  new_df[,2] <- as.numeric(og_df[,'y']) + delta_df[,'delta_y']
  new_df[,3] <- og_df[,3]
  new_df[,4] <- og_df[,4]
  new_df[,5] <- og_df[,5]
  colnames(new_df) <- c('x', 'y', 'subjects', 'incubation', 'recovery')
  new_df
}

  if(is.null(baddies)){
    return(new_df)
  }else{
  new_step <- get_step_length(new_df[i,], x_dim, y_dim)
  new_direction <- get_step_direction(new_step)
  new_change <- get_delts(new_step, new_direction)
  new_df <- update_coords(og_df[baddies,], new_change)
  full_df <- rbind(new_df, og_df[-baddies,])
  return(coord_checker(og_df, full_df, x_dim, y_dim))
  }
}

#Function for deciding if infections occurred
infection_status <- function(state, mean_crit_dist){

  current_infected <- state[state[,3]=='I' & state[,4] <= 0,]
  current_suceptable <- state[state[,3]=='S',]
  current_removed <- state[state[,3]=='R',]
  current_non_contagious <- state[state[,3]=='I' & state[,4] > 0,]

  #store new infection status
  newly_infected <- NULL

  #Check to see which susceptible subjects get infected
  #For loop
  if(nrow(current_suceptable) != 0 & nrow(current_infected) != 0){
  for(i in 1:nrow(current_infected)){
    for(j in 1:nrow(current_suceptable)){
      x_i <- as.numeric(current_infected[i, 'x'])
      y_i <- as.numeric(current_infected[i, 'y'])
      x_s <- as.numeric(current_suceptable[j, 'x'])
      y_s <- as.numeric(current_suceptable[j, 'y'])

      crucial_distance <- rexp(1, 1/mean_crit_dist)

      d <- sqrt((x_i - x_s)^2 + (y_i - y_s)^2)
      if(d <= crucial_distance){
        newly_infected <- c(newly_infected, j)
      }

    }
  }

  current_suceptable[newly_infected, 3] <- 'I'
  #Give incubation time
  current_suceptable[newly_infected, 4] <- rpois(n=length(newly_infected), lambda = 5)
  #Give recovery time
  current_suceptable[newly_infected, 5] <- rpois(n=length(newly_infected), lambda = 14)

  new_state <- rbind(current_infected, current_suceptable, current_removed, current_non_contagious)
  return(new_state)
  }else{return(state)}
}

#simulation function
simulator <- function(days, num_subjects, x_dim, y_dim, I_0, inf_dist){
  initial_data <- starting_vals(num_subjects = num_subjects, x_dim=x_dim, y_dim=y_dim, I_0=I_0)
  end_day <- list()
  num_cases <- rep(NA, days)
  num_cases <- c(I_0, num_cases)
  num_new_cases <- rep(NA, days)

  for(i in 1:days){

    if(i == 1){
      step_size <- get_step_length(initial_data, x_dim, y_dim)
      direction <- get_step_direction(step_size)
      change <- get_delts(step_size, direction)

      updated_positions <- update_coords(initial_data, change)
      checked_positions <- coord_checker(og_df=initial_data, new_df=updated_positions, x_dim=x_dim, y_dim=y_dim)
      new_status <- infection_status(checked_positions, inf_dist)

      #recovery time code
      not_recovered <- new_status[as.numeric(new_status[, 5]) > 0, ]
      new_time <- as.numeric(not_recovered[, 5]) - 1
      not_recovered[, 5] <- new_time
      infected_0 <- new_status[as.numeric(new_status[, 5]) == 0, ]
      infected_0[, 3] <- rep('R', nrow(infected_0))
      not_infected = new_status[as.numeric(new_status[, 5]) < 0, ]
      new_status = rbind(not_recovered, infected_0, not_infected)

      #incubation time code
      infected_people <- new_status[new_status[,3] == 'I',]
      new_inc_time <- as.numeric(infected_people[,4]) - 1
      infected_people[,4] <- new_inc_time
      other_people <- new_status[new_status[,3] != 'I',]
      new_status <- rbind(infected_people, other_people)

      end_day[[i]] <- new_status

      #Calculate number of new cases
      num_cases[i+1] <- sum(new_status[,3]=='I')
      num_new_cases[i] <- sum(new_status[,3]=='I') - I_0

    }

    else{
      step_size <- get_step_length(initial_data, x_dim, y_dim)
      direction <- get_step_direction(step_size)
      change <- get_delts(step_size, direction)

      updated_positions <- update_coords(end_day[[i-1]], change)
      checked_positions <- coord_checker(og_df=end_day[[i-1]], new_df=updated_positions, x_dim=x_dim, y_dim=y_dim)
      new_status <- infection_status(checked_positions, inf_dist)

      not_recovered <- new_status[as.numeric(new_status[, 5]) > 0, ]
      new_time <- as.numeric(not_recovered[, 5]) - 1
      not_recovered[, 5] <- new_time
      infected_0 <- new_status[as.numeric(new_status[, 5]) == 0, ]
      infected_0[, 3] <- rep('R', nrow(infected_0))
      not_infected = new_status[as.numeric(new_status[, 5]) < 0, ]
      new_status = rbind(not_recovered, infected_0, not_infected)

      #incubation time code
      infected_people <- new_status[new_status[,3] == 'I',]
      new_inc_time <- as.numeric(infected_people[,4]) - 1
      infected_people[,4] <- new_inc_time
      other_people <- new_status[new_status[,3] != 'I',]
      new_status <- rbind(infected_people, other_people)

      end_day[[i]] <- new_status

      #Calculate number of new cases
      num_cases[i+1] <- sum(new_status[,3]=='I')
      num_new_cases[i] <- sum(new_status[,3]=='I') - num_cases[i-1]

    }
  }
  return(num_cases)
}

#Testing
n=50
y <- simulator(n,100,10,10,2,1)
x <- 0:n

ggplot() +
  geom_line(aes(x=x, y=y)) +
  xlab('Day') +
  ylab('Number of Cases') +
  ggtitle('Number of Covid-19 cases over time')
```
